Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-07-15 Referee1_Response.tex

\noi
%We thank the referee for this suggestion. 
Thank you for this suggestion. 
You are right.
There are cases when generating scenarios is expensive in the stochastic optimization setting as well. 
We changed the introduction based on your, the associate editor and the other reviewer's comments. 
We specifically added this as a motivation. 
The motivation now appears in the first full paragraph on page 4 of the paper. 
\medskip

\begin{itemize}
\item[] \textit{However, perhaps more often, sampling the i.i.d. realizations is computationally cheap compared to solving the resulting SAA instances in stochastic optimization. 
In this latter case, we have a choice between simply increasing the number of batches in a non-overlapping method or using the overlapping batch means approach (with some degree of overlap) in order to reduce variance.
Based on the paper's results, perhaps the computational effort required to achieve a level of variance reduction should guide this choice. 
The authors do point to the fact that warm starts in a stochastic
optimization algorithm can be exploited much more easily when using overlapping batch means than when not using overlapping batches. 
Some attempt to explore this tradeoff would be valuable. 
The computational results use sorting (for the news vendor problem), a dynamic programming algorithm (for a stochastic knapsack model), and a decomposition method (for two-stage stochastic linear programs). 
While I am not suggesting that this exploration needs to be made for all three types of algorithms, even exploring it in the case of the sorting algorithm (where some algorithms are well suited for almost-sorted lists) could provide valuable insight.}
\end{itemize}

\noi 
We studied computational efficiency of the OMRP variance estimators and demonstrated efficiency on the newsvendor problem using insertion sort. 
The results of this study are presented in a new section, \S 5.4 (Computational Efficiency).
%We studied the computational efficiency of the OMRP estimators, based on your and the associate editor's comments, and provided a summary of our results in a new section, Section 5.4 (Computational Efficiency). 
We first approached efficiency by the reciprocal of variance multiplied by the computation time, per the associate editor's suggestion. 
Maximizing computational efficiency defined in this manner provided guidelines on selecting the degree of overlap. 
We then looked at how OMRP compared to increasing the number batches in MRP to reach an equivalent variance level.
We examined computational efficiency theoretically, assuming a functional form of the solution time per batch, and also empirically. 
For empirical evaluation, we used the newsvendor problem with insertion sort, based on your suggestions; thank you. 
We found that when effective warmstarting is available, overlapping can lead to more efficient estimators than (nonverlapping) MRP, both when the direct nonoverlapping counterpart is used or when the number of batches in MRP is increased to match the variance of the overlapping counterpart.  
%All these results are presented in Section 5.4 (Computational Efficiency). 
%Thank you for this suggestion. 
We believe the study of computational efficiency significantly improved the paper. 
Thank you.\smallskip

@@

\noindent 
Done.
\medskip

\begin{itemize}
\item[5.] \textit{p4, l34: Suggest: once the data are generated through a simulation, they \ldots}
\end{itemize}

\noindent 
Done. 
\smallskip

\noi 
{\it Note:} This sentence now appears in the second paragraph of \S 3 (Overlapping Multiple Replications Procedure) as a result of restructuring the introduction based on the comments of the associate editor and the other referee. 
\medskip

@@

\begin{itemize}
\item[6.] \textit{p4, l42: Suggest: Partial overlap results in fewer batches, hence fewer optimization \ldots}
\end{itemize}

\noindent 
We revised this section in order to improve exposition. 
We used `smaller number of batches' based on the comments of the other referee in the new sentence. 
\medskip

\noi 
{\it Note:} This section now appears in the second paragraph of \S 3 (Overlapping Multiple Replications Procedure) as a result of restructuring the introduction based on the comments of the associate editor and the other referee. 
\medskip

@@

\begin{itemize}
\item[7.] \textit{p6, l32: Suggest: The degrees \ldots make \ldots}
\end{itemize}

\noindent 
We now use denominator instead of degrees of freedom. 
Therefore, we used `\ldots denominator makes \ldots '
\medskip

@@

\begin{itemize}
\item[16.] \textit{p17, l55: Perhaps add a comment on why DB1 may not experience the asymptotic variance reduction? 
(Is it because it may lack a unique solution?)}
\end{itemize}

\noindent
Done.
% xxx Yes, it is partially because we are not sure if it has a unique optimal solution as DB1 has never been solved before. 
%Even if it has been solved, the variance reduction is shown asymptotically but the computations are only for small sample sizes. 
%We reworded this to reflect these points.
%Thank you for this suggestion. 
\medskip

\begin{itemize}
\item[17.] \textit{p18, figures: The figures are a bit small and hence can be hard to read without enlarging them (but perhaps this will be taken care of at production).}
\end{itemize}

\noindent 
We split these two figures into three and enlarged each graph. 
We also enlarged the legends in these graphs to improve readability.
All the new graphs added to the paper in \S 5.4 (Computational Efficiency) are also larger, with legible legends. 
\medskip

@@

\begin{itemize}
\item[18.] \textit{p18, l57: The authors speak of coverage probabilities for DB1, but how can this be done without knowing an optimal solution to DB1?}
\end{itemize}

\noindent 
We used a Latin hypercube sample of size 50,000 estimate $\mu_{\hat{x}}$ for DB1, which was then used to estimate the coverage probability of the OMRP confidence intervals for this test problem.
We explain this at the end of \S 5.1 (Test Problems). 
We also changed the wording in \S 5.3 (Results of Experiments) to make it clear that the coverage probability is estimated for DB1. 
\medskip

\begin{itemize}
\item[19.] \textit{p19: It may be possible to expand on the last paragraph of \S 5, in giving guidance on selecting the degree of overlap, depending on the authors' response to comment M1 above.}
\end{itemize}

### 2014-07-15 Referee2_Response.tex

%%%%%%%%%%%%%%%%%
%% Need to be careful with the minor changes. Will these be affected?
%% Need to CHECK AT THE VERY END... FOR ANY CHANGES on TOP OF these...
%%    - GB
%%%%%%%%%%%%%%%%%

\noi  
{\large \bf General Comments:}
\medskip

\begin{itemize}
\item[1.] \textit{The main contributions of the paper need to be stated more clearly and succinctly.}
\end{itemize}

\noindent 
You are right. 
We changed the introduction to clearly state the main contributions of the paper.  
Please see page four of the revised manuscript. 
We also clarified the motivation for the work, included new contributions based on the study of computational efficiency (contribution (iv)), and provided a comparison with the proceedings version of the paper (based on the editor-in-chief's suggestion).
%We believe the presentation of the paper significantly improved.  
%Thank you.
\medskip

@@

\begin{itemize}
\item[2.] \textit{It is not necessary to go into such detail in the OBM review.}
\end{itemize}

\smallskip

\noindent 
We shortened and streamlined the OBM review, focusing on the main results used in the paper, by combining two previous sections---\S 2.1 (Overlapping Batch Means) and \S 3.1 (Variably Overlapping Batches)---along with other explanations on OBM that appeared later in the paper. 
The combined review is in the background section, \S 2.1 (Batch Means). 
To improve readability, we created two subsections: \S 2.1.1 (Nonoverlapping Batch Means) and \S 2.1.2 (Overlapping Batch Means). 
While we removed a number of equations and shortened the text, we kept some explanations that are necessary for a reader who is not familiar with OBM (e.g., readers with an optimization focus) to follow the main ideas. 
These basic explanations appear, for instance, in the first paragraph of \S 2.1.2 (Overlapping Batch Means).  
We hope to be able to reach a broader audience this way. 
As a result of these changes, 
\begin{itemize}
	\item[(i)]  instead of former equations (2) and (5), both on overlapping batch means variance estimator, we have only one equation (2);

@@

\bigskip

\noi  
{\large \bf Specific Comments:}
\medskip

\begin{itemize}
\item[] \textit{Page 1, Line 32 (and elsewhere): I believe that you ``apply the OBM method to stochastic programming'' rather than extend it.}
\end{itemize}

\noi
Changed.  
We also went through the paper to correct other occurrences. 
In particular, this was also used in \S 6 (Conclusions). 
We changed the first sentence of \S 6 (Conclusions) to use the word `apply' instead of `extend'. 
\medskip

@@

\begin{itemize}
\item[] \textit{2, 5: ``never before been studied . . . context.''}
\end{itemize}

\noi
Changed.
\medskip

\begin{itemize}
\item[] \textit{2, 7: By ``same benefits'', you mean moderate bias and comparatively low variance, right?}
\end{itemize}

\noi
Yes. We reworded this sentence. 
We also went through the paper and made the context clear whenever phrases like `same benefits', `desired statistical properties', etc.\ were used.  
\medskip

@@

\noi
We made the change to the first sentence above. 
The first sentence now appears as part of contribution (ii) and the second sentence as part of contribution (iii) on page 4 of the revised manuscript. 
We decided not to include `also' in the second sentence as it starts a new set of contributions. 
\medskip

\begin{itemize}
\item[] \textit{2, 28: ``Our experiments indicate that the use of overlapping estimators results in gap estimators achieving lower variance than their nonoverlapping counterparts, along with comparable bias and interval coverage --- for both small- and large-sample cases.}
\end{itemize}

\noi
Done.
\medskip

\noi
{\it Note:} 
This sentence now appears as part of contribution (iii) on page 4 of the revised manuscript.
\medskip

@@

\noi
Deleted.
\medskip

\begin{itemize}
\item[] \textit{4, 17: ``interest is that of estimating''}
\end{itemize}

\noi
Done.  
\medskip

\noi
{\it Note:}
 As a result of the revised introduction---aimed to clearly state the contributions of the paper---we moved the technical discussion on using overlapping batches in simulation output analysis versus for assessing solution quality in stochastic optimization to \S 3 (Overlapping Multiple Replications Procedure). 
Consequently, this sentence now appears in \S 3. 
\medskip

@@

\begin{itemize}
\item[] \textit{4, 20--27: Even though I believe everything, these lines are a bit awkward and confusing.}
\end{itemize}

\noi
You are right. 
We removed the comment on the central limit theorem on the minimized sample means as it is not relevant to the paper. 
We also reworded these sentences to improve clarity. 
\bigskip

\noi
{\it Note:}
As mentioned in the previous comment, the revised sentences now appear in \S 3 (Overlapping Multiple Replications Procedure). 
\medskip

@@

\begin{itemize}
\item[] \textit{4, 31: What are the desired statistical properties?}
\end{itemize}

\noi
Thanks for pointing out the ambiguity here. 
We reworded this sentence, divided it into two sentences to improve readability, and added further explanations to remove any ambiguities. 
\bigskip

\noi
{\it Note:}
The revised sentences now appear in \S 3 (Overlapping Multiple Replications Procedure). 
\medskip

@@

\noi
Done.
We also changed the sentence to make it clear what the ``difference'' is referring to. 
\medskip

\noi
{\it Note:}
This sentence now appears in \S 3 (Overlapping Multiple Replications Procedure). 
\medskip

\begin{itemize}
\item[] \textit{4, 36: ``due to the need to solve sampling problem"}
\end{itemize}

\noi
Changed it to ``\dots because we need to solve \ldots '' to improve readability.  
\medskip

@@

\noi
{\it Note:}
This sentence now appears in \S 3 (Overlapping Multiple Replications Procedure). 
\medskip

\begin{itemize}
\item[] \textit{4, 42: ``Partial overlapping results in a smaller number of batches; hence, fewer optimization''}
\end{itemize}

\noi
We revised this section in order to improve exposition. 
We used `smaller number of batches' based on your suggestion in the revised version. 
\medskip

\noi 
{\it Note:} This sentence now appears in the second paragraph of \S 3 (Overlapping Multiple Replications Procedure). 
\medskip

@@

\begin{itemize}
\item[] \textit{4, 45: The sentence ``Second, \ldots '' needs some clarification, since it states a conclusion that hasn't so far been justified.}
\end{itemize}

\noi
You are right. 
While we removed the sentence that started with `Second' in an effort to improve readability, we referred to the computational efficiency study added to the paper to justify this remark.   
\medskip

@@

\begin{itemize}
\item[] \textit{4, 54 (and elsewhere): When you refer to ``variable amounts of overlap'', that is really nothing new. Certainly, any overlapping estimator will have end effects in which the overlap amounts differ from those in the ``middle'' region of the data.}
\end{itemize}

\noi
%You are right.
We meant to say that the degree of overlap is parametrized and this parameter can be chosen by the user. 
We reworded this sentence to reflect this. 
We also went through the paper, removed other occurrences, and avoided using this term in the new text added. 
\medskip

\begin{itemize}
\item[] \textit{4, 57 (and elsewhere): ``classical case'' refers to nonoverlapping batch means, right?}
\end{itemize}

\noi
No. By `classical case', we meant the use of overlapping batches in simulation output analysis (as opposed to assessment of solution quality in stochastic programming). 
We reworded this sentence to remove ambiguity. 
Based on your earlier comments, we avoided the term `variance reduction' in the revised sentence.
We also went through the paper and removed several uses of the word `classical' that referred to nonoverlapping batches to reduce ambiguity.
%(as these were referring to nonoverlapping batches). 
%In the current version of the paper, `classical case' is used to only indicate  the use of overlapping batches in simulation output analysis. 
\medskip

@@

\begin{itemize}
\item[] \textit{5, 14: ``realization of the univariate stochastic''}
\end{itemize}

\noi
Done.  
\medskip

\begin{itemize}
\item[] \textit{5, 35 (and elsewhere): Careful! This isn't quite the sample variance! This is actually the sample variance multiplied by $m/n$.}
\end{itemize}

\noi
Thanks for catching this. 
We corrected this sentence. 
Please note that we split this sentence into two to avoid long sentences and to improve exposition. 
We also checked for other uses of the term `sample variance' throughout the paper and corrected these as needed.
Thank you.  
\medskip

@@

\begin{itemize}
\item[] \textit{5, 43: You could have introduced the notation ``CI'' earlier.}
\end{itemize}

\noi
You are right. 
We now introduce this abbreviation the first time we use the term `confidence interval' in the introduction. 
From this point on, we use the abbreviated form `CI' throughout the paper, except for the title of \S 4.5 (Asymptotic Validity of the Confidence Intervals). 
\medskip

@@

\begin{itemize}
\item[] \textit{6, 13: ``corresponding overlapping batch means''}
\end{itemize}

\noi
Done.  
We removed the word `corresponding' as part of the rewrite.
\medskip

\begin{itemize}
\item[] \textit{6, 29: Note that you use ``Note that'' too often.}
\end{itemize}

\noi
We went through the paper and removed most of the ``note that'' phrases. 
We decided to leave a couple as is (as ``notice that'') because they highlight important characteristics.  
\medskip

@@

\begin{itemize}
\item[] \textit{6, 30: Explicitly say what the d.f.\ in (2) is.}
\end{itemize}

\noi
We changed this sentence and directly referred to the denominator term.
We decided to avoid the degrees of freedom here. 
This is because we wanted to focus on the degrees of freedom for confidence interval construction, which is discussed in both the theoretical and computational results.
\medskip

@@

\noi 
{\it Note:} This section has been restructured based on your comments. 
Equation (2) now gives the general overlapping variance estimator with the parametrized amount of overlap. 
\medskip

\begin{itemize}
\item[] \textit{7, 46: Does $\gamma$ need to be an integer? 
You seem to be using it that way, so maybe state the requirement.}
\end{itemize}

\noi
Yes; we explicitly stated this in a sentence after we introduce $\gamma$.
\medskip

\noi 
{\it Note:} This sentence now appears in \S 2.1.2 (Overlapping Batch Means). 
\medskip

@@

\begin{itemize}
\item[] \textit{8, 4 (and elsewhere): Define notations such as floor and ceiling functions.}
\end{itemize}

\noi
Done. 
%We introduced the floor function before its first use, and we introduced the ceiling function right after its first use.
\smallskip

\noi 
{\it Note:} Both definitions appear in \S 2.1.2 (Overlapping Batch Means).
\medskip

@@

\begin{itemize}
\item[] \textit{8, 11: In addition to the fact that I don't think ``variance reduction'' is quite the right term, the phrase ``variance reduction in this estimator'' is awkward.}
\end{itemize}

\noi
Reworded.\smallskip

\noi 
{\it Note:} This sentence now appears in \S 2.1.2 (Overlapping Batch Means). 
\medskip

@@

\begin{itemize}
\item[] \textit{8, 53--57: The various OBM references should have been discussed earlier.}
\end{itemize}

\noi
You are right. 
We moved these references---in particular, Damerdji (1994, 1995)---to \S 2.1.2 (Overlapping Batch Means) and added a brief discussion.  
\medskip

@@

\begin{itemize}
\item[] \textit{12, 52: ``$n_l VD_l$ has been studied''}
\end{itemize}

\noi
Done.
\medskip

\begin{itemize}
\item[] \textit{13, 12 (and elsewhere): I may have missed this, but has the notation $\sigma_{\hat{x}}$ been established yet?}
\end{itemize}

\noi
This notation was defined before Theorem 1. 
We moved the definitions of $\sigma_{\hat{x}}$ and $\mu_{\hat{x}}$ to the end of \S 4.2 (Nonoptimized Counterparts) to improve clarity. 
\medskip

@@

\begin{itemize}
\item[] \textit{15, 25: Do you discuss $t_{d_l,\alpha}$ at all in Theorem 3?}
\end{itemize}

\noi
No, we did not.   
We added explanations in the proof Theorem 3, discussing how $t_{d_l,\alpha}$ factors in the proof. 
\medskip

\begin{itemize}
\item[] \textit{18: Your figures are very difficult to read, even with the pdf file.}
\end{itemize}

\noi
We split these two figures into three and enlarged each graph. 
We also enlarged the legends in these graphs to improve readability.
All the new graphs added to the paper in \S 5.4 (Computational Efficiency) are also larger, with legible legends. 
\medskip

### 2014-07-15 AE_Reponse.tex
Crossing the t's, dotting the i's

\noi 
Thank you for this suggestion. 
It is indeed important to study computational efficiency in this context. 
We created a new section, \S 5.4 (Computational Efficiency), to study the computational efficiency of overlapping estimators in the stochastic optimization setting. 
%When an algorithm has warm-starting capability, as the degree of overlapping increases,  solution time per batch could decrease; as a result, computational efficiency gains could be observed.  
We discussed efficiency for a given functional form of the solution time per batch. 
%Depending on the effectiveness of warm-starting, maximizing computational efficiency provides guidelines on selecting the degree of overlapping. 
%\medskip 
In addition, based on the comments of Referee 1, we studied efficiency of overlapping estimators compared to increasing the number of batches in MRP to obtain (nonoverlapping) estimators with equivalently lower variances. 
Our results indicate that when warm-starting is effective, overlapping is computationally more efficient.
Even in cases where the effectiveness of warm-starting is quite low, computational efficiency can be observed with a low degree of overlap (e.g., when approximately 10\% of observations are overlapped).\bigskip

@@

\noi 
We established improved computational efficiency on the newsvendor problem using insertion sort.
We also (i) changed Section 1 (Introduction) to clearly motivate the study and argue usefulness of the results, paying specific attention to efficiency gains; (ii) updated Section 6 (Conclusions) to summarize these new results, again paying attention to computational efficiency; (iii) provided a study of computational efficiency gains of overlapping in the context of assessing solution quality in stochastic programming in the new \S 5.4 (Computational Efficiency); and finally (iv) updated the abstract.  
%(Please see our response to the previous comment.)
\medskip

@@

>>>>

\noi 
Thank you for this suggestion. 
It is indeed important to study computational efficiency in this context. 
We created a new section, \S 5.4 (Computational Efficiency), to study the computational efficiency of overlapping estimators in the stochastic optimization setting. 
%When an algorithm has warmstarting capability, as the degree of overlapping increases,  solution time per batch could decrease; as a result, computational efficiency gains could be observed.  
We discussed efficiency for a given functional form of the solution time per batch. 
%Depending on the effectiveness of warmstarting, maximizing computational efficiency provides guidelines on selecting the degree of overlapping. 
%\medskip 
In addition, based on the comments of Referee 1, we studied efficiency of overlapping estimators compared to increasing the number of batches in MRP to obtain (nonoverlapping) estimators with equivalently lower variances. 
Our results indicate that when warmstarting is effective, overlapping is computationally more efficient.
Even in cases where the effectiveness of warmstarting is quite low, computational efficiency can be observed with a low degree of overlap (e.g., when approximately 10\% of observations are overlapped).\bigskip

@@

\noi 
We established improved computational efficiency on the newsvendor problem using insertion sort.
We also (i) changed \S 1 (Introduction) to clearly motivate the study and argue usefulness of the results, paying specific attention to efficiency gains; (ii) updated \S 6 (Conclusions) to summarize these new results, again paying attention to computational efficiency; (iii) provided a study of computational efficiency gains of overlapping in the context of assessing solution quality in stochastic programming in the new \S 5.4 (Computational Efficiency); and finally (iv) updated the abstract.  
%(Please see our response to the previous comment.)
\medskip

@@

\noi
{\it Note:}
We moved this sentence from \S 1 (Introduction)---to clearly state the contributions of the paper---to \S 3 (Overlapping Multiple Replications Procedure), where we provided more technical details. 
\medskip
